---
title: "Poster figure -- Wall Table 1 reproduction (Student L)"
format: html
---

You're filling in the **comparison of Table 1 from Wall (2004)** with the
queen-contiguity replication. The poster comment asks for a side-by-side
comparison of the IID, SAR, and CAR fits. Student C will then add the
Hausdorff-weight versions to the same table.

The `ReplicatingWall.qmd` file already fits all three models -- your job
is to assemble the comparison table and a short visual.

## Setup (same as the replication file)

```{r}
#| message: false
library(sf)
library(spatialreg)   # brings in `us_states` from spData; provides spautolm
library(dplyr)
library(ggplot2)
library(spdep)
library(tibble)
library(tidyr)
library(patchwork)
library(janitor)
```

## Read data + set up geometry + queen weights (copied from the replication file)

```{r}
wall_data <- read.csv("data/wall_2004_dataset.csv")

wall_data <- wall_data %>%
  filter(!(name %in% c("District of Columbia", "Alaska", "Hawaii", "USA")))

wall_data$perc_elig_2 <- wall_data$perc_elig^2

wall_sp <- us_states %>%
  inner_join(wall_data, join_by(NAME == name))

wall_sp_ordered <- wall_sp %>% arrange(GEOID) %>%
  mutate(GEOID = seq(from = 1, to = 48, by = 1))

# queen-contiguity weights -- Wall's original spec
nb_us <- poly2nb(wall_sp_ordered, queen = TRUE)
lw_queen <- nb2listw(nb_us, style = "W")
```

## Fit the three models (copied from the replication file)

```{r}
fit_iid <- lm(verbal ~ perc_elig + perc_elig_2, data = wall_data)

summary(fit_iid)

fit_SAR <- spautolm(verbal ~ perc_elig + perc_elig_2,
                    data   = wall_sp_ordered,
                    listw  = lw_queen,
                    family = "SAR")
summary(fit_SAR)
fit_CAR <- spautolm(verbal ~ perc_elig + perc_elig_2,
                    data   = wall_sp_ordered,
                    listw  = lw_queen,
                    family = "CAR")

summary(fit_CAR)
```

## YOUR TASK 1 -- assemble the comparison table

We want one row per model with columns: `intercept`, `perc_elig`,
`perc_elig_2`, `spatial_param` (lambda for SAR/CAR, NA for IID), and
`logLik`.

```{r}
# Helper to pull the right pieces out of a fit.
extract_row <- function(model_name, fit) {
  if (inherits(fit, "lm")) {
    cf <- coef(fit)
    spatial <- NA_real_
  } else {
    # spautolm objects: coef() returns spatial param + betas
    cf_full <- coef(fit)
    spatial <- cf_full["lambda"]   # SAR / CAR spatial parameter
    cf <- cf_full[c("(Intercept)", "perc_elig", "perc_elig_2")]
  }
  
  print(cf)
  tibble(
    Model         = model_name,
    intercept     = cf["(Intercept)"],
    perc_elig     = cf["perc_elig"],
    perc_elig_2   = cf["perc_elig_2"],
    spatial_param = as.numeric(spatial),
    logLik        = as.numeric(logLik(fit))
  )
}

# TODO: build `comparison_tbl` by binding the three rows together.
#   Use bind_rows() and call extract_row() once per fit.
#   Order: IID, SAR, CAR.

comparison_tbl_iid <- extract_row(model_name = "iid", fit = fit_iid)

comparison_tbl_SAR <- extract_row(model_name = "SAR", fit = fit_SAR)

comparison_tbl_CAR <- extract_row(model_name = "CAR", fit = fit_CAR)

comparison_tbl = bind_rows(comparison_tbl_iid, comparison_tbl_SAR, comparison_tbl_CAR)

comparison_tbl %>% rename("Model" = Model, 
                          "Intercept" = intercept, 
                          "Beta1" = perc_elig, 
                          "Beta2" = perc_elig_2, 
                          "Spatial Parameter" = spatial_param,
                          "Log likelihood" = logLik) %>% 
  t() %>% 
  data.frame() %>% 
  row_to_names(row_number = 1)



```

## YOUR TASK 2 -- save it for student C to extend

C will read this back in and append two more rows (SAR + CAR with Hausdorff
weights). Don't change the column names or the file path.

```{r}
#saveRDS(comparison_tbl, "data/queen_fits_table.rds")
```

## YOUR TASK 3 -- a small comparison visualization

A helpful poster visual is a bar chart of `logLik` across the three models,
since better-fitting spatial models should give a higher log-likelihood than
the IID baseline. The IID bar gives the visual baseline.

```{r}
#TODO: bar chart of logLik by Model.

ggplot(comparison_tbl, aes(x = Model, y = logLik, fill = Model)) + 
  geom_col(width = 0.6) + 
  geom_text(aes(label = round(logLik, 1)), vjust = 1.4, size = 3) + 
  labs(title = "Log-likelihood by model (queen contiguity)",
     x = NULL, y = "log L") + 
  theme_minimal() + theme(legend.position = "none")
#
  
#Save as p_loglik.

p_loglik <- ggplot(comparison_tbl, aes(x = Model, y = logLik, fill = Model))
  # TODO
```

## YOUR TASK 4 -- map the SAR fitted values

Pair the bar chart with the spatial fitted values from the SAR model so the
poster reader sees both the numerical and the spatial story.

```{r}
wall_sp_ordered$.fitted_sar <- fitted(fit_SAR)

# TODO: choropleth of .fitted_sar.


ggplot(wall_sp_ordered, aes(fill = .fitted_sar)) +
  geom_sf() +
  scale_fill_gradientn(
    colours = grey.colors(4, start = 0.9, end = 0),
    breaks = c(0, 506, 530, 565),
    name = "Verbal SAT") +
  labs(title = "SAR fitted values (queen contiguity)") +
  theme_minimal()

# Save as p_map.

p_map <- ggplot(wall_sp_ordered, aes(fill = .fitted_sar)) +
  geom_sf()
  # TODO
```

## Combine and save

```{r}
panel <- (p_loglik | p_map) +
  plot_annotation(
    title = "Wall (2004) Table 1 -- queen-contiguity replication",
    theme = theme(plot.title = element_text(face = "bold", size = 14))
  )

panel

ggsave("figures/wall_queen_comparison.png",
       panel, width = 11, height = 5, dpi = 200, bg = "white")
```

## Stretch -- side-by-side residual maps

If you have time: replicate `wall_sp_ordered` with three columns
(`.resid_iid`, `.resid_sar`, `.resid_car`), pivot longer, and facet by
model:

```{r}
#| eval: false
wall_sp_ordered$.resid_iid <- residuals(fit_iid)
wall_sp_ordered$.resid_sar <- residuals(fit_SAR)
wall_sp_ordered$.resid_car <- residuals(fit_CAR)

resid_long <- wall_sp_ordered %>%
  pivot_longer(starts_with(".resid_"),
               names_to = "model", values_to = "resid",
               names_prefix = "\\.resid_")

ggplot(resid_long) +
  geom_sf(aes(fill = resid)) +
  scale_fill_gradient2(low = "#2166ac", mid = "white", high = "#b2182b",
                       midpoint = 0, name = "Residual") +
  facet_wrap(~ model, ncol = 3) +
  theme_minimal()
```
